---
title: "APS Odd Semester 2024 Coding Problem Set-2"
editor_options:
  chunk_output_type: console
output:
  word_document: default
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r}
library(ggplot2)
library(dplyr)
library(corrplot)

```



```{r}
# Load the house price d
hData = read.csv('houseprices.csv', header = TRUE, stringsAsFactors = FALSE, na.strings = c("", "NA", "Not Available", "not available"))
str(hData)
```


```{r}
na.omit(hData)

```

```{r}
# Convert 'locality', 'facing' and 'parking' columns to factors
categorical_cols = c('locality', 'facing', 'parking')
hData[categorical_cols] = lapply(hData[categorical_cols], as.factor)
str(hData)
```

```{r}
# Continuous columns
continuous_cols = setdiff(colnames(hData), categorical_cols)
```

```{r}
# Plot percentage of NAs in each column of the data frame
hData_NA = setNames(stack(sapply(hData, function(x){(sum(is.na(x))/length(x))*100}))[2:1], c('Feature','Value'))
p = ggplot(data = hData_NA, aes(x = Feature, y = Value)) +
  geom_bar(stat = 'identity', fill = 'steelblue', width = 0.3) +
  theme(text = element_text(size = 14, face = 'bold'),
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  xlab('') + ylab('Percentage') +
  ggtitle('Percentage of NAs across all features')
p
```

#can see that there are none values in facing and parking

```{r}
# Add NA as a factor level for categorical columns 
hData[categorical_cols] = lapply(hData[categorical_cols], addNA)
str(hData)
```



```{r}
# Scatter Plot: Rent vs Area, colored by BHK
ggplot(hData, aes(x = area, y = rent, color = factor(BHK))) +
  geom_point(alpha = 0.7) +
  labs(title = "Scatter Plot of Rent vs Area (Colored by BHK)",
       x = "Area (sqft)", y = "Rent (INR)") +
  theme_minimal()


#1-As the area increases, rent generally tends to rise.
#2-Properties with higher BHK values are clustered at higher rents, indicating that #larger properties (more rooms) are priced higher.



```


```{r}
# correlation matrix
numeric_cols <- hData[, c("area", "rent", "price_per_sqft", "BHK", "bathrooms")]

# Compute the correlation matrix
cor_matrix <- cor(numeric_cols, use = "complete.obs")
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45, addCoef.col = "black")

#Correlation Analysis - 
 #1- (r=0.79): Rent tends to increase as area increases.
 
#2- Price per Sqft and Bathrooms (0.080264)-
#very weak positive correlation btw price_per_sqft and bathrooms. The number of bathrooms has little to no effect on the price per square foot.

#3- Area and Bathrooms (0.812822):
#There is a strong positive correlation between area and bathrooms. Larger properties tend to have more bathrooms.

#4-Price per Sqft and Bathrooms (0.080264):

#There is a very weak positive correlation between price_per_sqft and bathrooms. The number of bathrooms has little to no effect on the price per square foot.


```


```{r}

mean_rent <- mean(hData$rent)
sd_rent <- sd(hData$rent)

# Plot the histogram with density and lines for mean and standard deviation
ggplot(hData, aes(x = rent)) +
  geom_histogram(aes(y = ..density..), bins = 100, fill = "green", color = "black", alpha = 0.7) +  # Histogram with outlines
  geom_density(color = "black", size = 1) +  # Add density curve
  geom_vline(aes(xintercept = mean_rent), color = "blue", linetype = "dashed", size = 1) +  # Mean line
  geom_vline(aes(xintercept = mean_rent + sd_rent), color = "red", linetype = "dotted", size = 1) +  # +1 Standard Deviation
  geom_vline(aes(xintercept = mean_rent - sd_rent), color = "red", linetype = "dotted", size = 1) +  # -1 Standard Deviation
  labs(title = "Distribution of Rent with Mean and Standard Deviation", x = "Rent", y = "Density") +
  
  # Add text label for mean rent on the graph
  annotate("text", x = mean_rent, y = 0.00002, label = paste("Mean =", round(mean_rent, 2)), color = "blue", angle = 90, vjust = -0.5, size = 4) +
  
  theme_minimal()

```

---
histogram of rent prices shows a right-skewed distribution, indicating that most properties have relatively lower rent prices, with a few high-rent outliers pulling the average up. The mean rent is approx 61,652, as represented by a dashed blue line on the graph.  This wide standard deviation reflects a large variation in rent values, with a spread from affordable to luxury properties 
---


```{r}
# Calculate and print summary statistics for Rent
rent_mean <- mean(hData$rent, na.rm = TRUE)
rent_sd <- sd(hData$rent, na.rm = TRUE)
rent_var <- var(hData$rent, na.rm = TRUE)
cat("Rent - Mean:", rent_mean, " | SD:", rent_sd, " | Variance:", rent_var, "\n")
```

```{r}
ggplot(hData, aes(x = parking, y = area)) +
  geom_boxplot() +
  labs(title = "Box Plot of Rent by Parking", x = "Parking", y = "Area") +
  theme_minimal()
```

```{r}
ggplot(hData, aes(x = locality)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(
    title = "Number of Houses by Locality", 
    x = "Locality", 
    y = "Number of Houses"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
```


```{r}
ggplot(hData, aes(x = locality, y = price_per_sqft, fill = locality)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8) +
  labs(
    title = "Distribution of Price per Sqft by Locality", 
    x = "Locality", 
    y = "Price per Sqft (INR)"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```


```{r}
# Box plot of Rent by Parking with colors
ggplot(hData, aes(x = as.factor(parking), y = rent, fill = as.factor(parking))) +
  geom_boxplot() +
  labs(title = "Rent Distribution by Parking Spaces",
       x = "Parking Spaces",
       y = "Rent (INR)") +
  scale_fill_brewer(palette = "Set3") +  # Choose a color palette
  theme_minimal() +
  theme(legend.title = element_blank())  # Remove legend title

```


