---
title: "Coding Assignment on dplyr and ggplot"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}
library(ggplot2)
library(dplyr)
```

#### Load the bodyfat dataset
```{r}
# Load the bodyfat dataset
file = 'bodyfat.csv'
df = read.csv(file, header = TRUE, row.names = 1 ,stringsAsFactors = FALSE)
str(df)
head(df, n = 2)
```

#### Create lists of categorical and continuous features
```{r}
# Create lists of categorical and continuous features
features = colnames(df)
categorical_features = c('Gender')
continuous_features = features[!(colnames(df) %in% categorical_features)]
```

#### Convert categorical columns to correct type
```{r}
# Convert categorical columns to correct type
df[categorical_features] = lapply(df[categorical_features], as.factor)
str(df)
```

#### Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
```{r}
# Summarizing sample data: mean, variance, standard deviation, median, percentiles, and quantiles
mean(df$Weight)
df$Weight=mean(df$Weight)#deviation vector
(df$Weight=mean(df$Weight))^2 #sq dev vector
mean((df$Weight-mean(df$Weight))^2) #sample
sqrt(mean((df$Weight-mean(df$Weight))^2)) #sample std deviation 
var((df$Weight)) #sample variance using inbuilt fucntion 
sd(df$Weight)
(df$Weight-mean(df$Weight))/sd(df$Weight)
scale(df$Weight)

```

#### Plot the distribution of males and females in the dataset
```{r}
# Plot the distribution of males and females in the dataset
p = ggplot(data = df) +
  geom_bar(aes(x = Gender),color='black' , fill= 'steelblue' ) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Gender',
       y = 'Frequency',
       title = 'Distribution of Genders')
p

```


#### Plot a component plot of weight values
```{r}
# Plot a component plot of weight values
p = ggplot(data = df) +
  geom_point(aes(x= 1:nrow(df),y=Weight)) +
  geom_hline(yintercept = mean(df$Weight),color='red')+
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```

```{r}
# Plot a component plot of weight values
p = ggplot(data = df) +
  geom_point(aes(x= 1:nrow(df),y=Weight-mean(Weight))) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Weight (kg)',
       title = 'Component Plot of Weights')
p
```


```{r}
p = ggplot(data = df) +
  geom_point(aes(x= 1:nrow(df),y=Bodyfat-mean(Bodyfat))) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Patient #',
       y = 'Bodyfat (kg)',
       title = 'Component Plot of Weights')
p
```

#### Plot a histogram of weight values
```{r}
# Plot a histogram of weight values
mean_weight = mean(df$Weight)
sd_weight = sd(df$Weight)
p = ggplot(data = df) +
  geom_histogram(aes(x=Weight,y=after_stat(count)),
  breaks = seq(mean_weight-3*sd_weight,
   mean_weight+3*sd_weight,by = 2.0),
   color = 'black',fill='steelblue', alpha = 0.4) +
  geom_vline(xintercept = mean_weight, color = 'red', linetype = 'dashed') +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Weight (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Weights')
p

```

#### Plot a histogram of height values
```{r}
# Plot a histogram of height values
mean_weight = mean(df$Bodyfat)
sd_weight = sd(df$Bodyfat)
p = ggplot(data = df) +
  geom_histogram(aes(x=Bodyfat,y=after_stat(count)),
  breaks = seq(mean_weight-3*sd_weight,
   mean_weight+3*sd_weight,by = 2.0),
   color = 'black',fill='steelblue', alpha = 0.4) +
  geom_vline(xintercept = mean_weight, color = 'red', linetype = 'dashed') +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
  axis.text = element_text(size = 12),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14),
  axis.title = element_text(size = 14, face = "bold"))+
  labs(x = 'Bodyfat (kg)',
       y = 'Frequency',
       title = 'Frequency Histogram of Bodyfat')
p
```

#### Plot a histogram of bodyfat values
```{r}
# Plot a histogram of bodyfat values


```

#### Plot a histogram of weights for both genders
```{r}
# Plot a histogram of weights for both genders

```

#### Plot a boxplot of bodyfat values
```{r}
# Plot a boxplot of bodyfat values
p <- ggplot(data = df) +
  geom_boxplot( aes(x=Gender , y = Bodyfat)) +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        axis.title = element_text(size = 14, face = "bold")) +
  labs(x = '',
       y = 'Bodyfat (kg)',
       title = 'Boxplot of Bodyfat')
p
```

#### Plot a boxplot of bodyfat values for both genders
```{r}
# Plot a boxplot of bodyfat values for both genders
p = ggplot(data = df) +
  geom_boxplot(aes(x=Gender,y = Bodyfat,fill=df$Weightlevel))+
   labs(x = 'Gender',
       y = 'Bodyfat',
       title = 'Box plot')
p
```

#### Introduce a new categorical column in the dataframe called 'Weightlevel' with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)
```{r}
# Introduce a new categorical column in the dataframe called weight level with the following levels: (1) low (< 25th percentile of weights) (2) medium (in the IQR of weights) (3) high (>75th percentile of weights)

# Calculate the 25th and 75th percentiles
q25 <- quantile(df$Weight, 0.25)
q75 <- quantile(df$Weight, 0.75)

# Introduce a new categorical column 'Weightlevel'
df <- df %>%
  mutate(Weightlevel = case_when(
    Weight < q25 ~ "low",
    Weight > q75 ~ "high"
  ))



```

#### Plot a boxplot of bodyfat values for both genders across the three Weightlevel groups
```{r}
# Plot a boxplot of bodyfat values for both genders both genders across the three Weightlevel groups
````

#### The 68-95-99 rule
```{r}
# The 68-95-99 rule

```


