---
title: "Conditional probability using simulation"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```


#### Load the bats data to a dataframe
```{r}
dBat = read.csv('bats.csv', header = TRUE,row.names=1 , stringsAsFactors = TRUE)
head(dBat)
```

##### Get the genes that are expressed for the 1st bat
```{r}
head(dBat,1)
true_values <- dBat[1, ][dBat[1, ] == TRUE]
print(true_values)
```

#### Probability that a random bat carries the Ebola virus
```{r}



```

#### Probability that gene-1 through gene-5 is expressed in a random bat
```{r}

```

#### Is expression of gene-1 through gene-5 dependent or independent of the bat carrying the Ebola virus
```{r}
# Subset the dataframe to those bats that carry the Ebola virus


# Conditional probability that gene-1 through gene-5 is expressed
# given that the bat carries the Ebola virus

```

#### Probability that a random bat carries the Ebola given that gene-3 through gene-5 is expressed
```{r}
setwd("C:/Users/Suyash/Downloads/archive (1)")
wnba=read.csv("wnba.csv")
wnba = wnba %>% rename(FG_perc=FG. , Three_PA=X3PA , Fifteen=X15.00)
print(names(wnba))
```

