---
title: "Quantifying uncertainty using simulation"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Sampling space and likelihood vector
```{r}
s = c('H', 'T')
p = c(0.5, 0.5)
```


#### Function to simulate one trial of the random experiment which in this case is tossing 10 fair coins at a time
```{r}
simTrial = function(){
  return(sample(s, 10, replace = TRUE, prob = p))
}
```

#### Replicate the trials of the random experiment a large number of times
```{r}
nsimulations = 10000
simulatedData = replicate(nsimulations, simTrial())
```

#### Function to check if event of interest has occured which in this case is getting exactly 3 heads
```{r}
checkEvent = function(data){
  return(sum(data == 'H') == 3)
}
```

#### Calculate the fraction of times the event of interest occured which is what we define as the approximate probability of the event E happening in the next trial of the random experiment
```{r}
#simulatedData
mean(apply(simulatedData, 2, checkEvent))
```
