---
title: "Intro to R"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#### Execute the following cells to load the libraries
```{r}
library(ggplot2)
library(dplyr)
```

#### Load the food texture dataset
```{r}
#file = 'http://openmv.net/file/food-texture.csv'
file = 'Data/food-texture.csv'
foodData = read.csv(file, header = TRUE, row.names = 1, stringsAsFactors = FALSE)
head(foodData, 2)
str(foodData)
```

#### Modify Crispy column to reflect high (0) or low (1) crispiness
```{r}
foodData = foodData %>% mutate(Crispylevel = ifelse(Crispy > 11, 'High', 'Low'))
```

##### Change Crispy and Crispylevel columns to factor (categorical) type
```{r}
# Continuous features -> Oil, Density, Hardness, Fracture
# Categorical features -> Crispy (8 levels 8 through 15) and Crispylevel (2 levels 0 and 1)
# Crispy is a categorical feature with an order
str(foodData)
categorical_cols = c('Crispy', 'Crispylevel')
foodData[categorical_cols] = lapply(foodData[categorical_cols], as.factor)
str(foodData)
```

#### Visualize the OilPercentage feature using a histogram
```{r}
```

#### Scatter plot between OilPercentage and Density
```{r}
```

#### Scatter plot between OilPercentage and Density color coded with Crispylevel
```{r}
```

#### Scatter plot between Density and Hardness
```{r}
```

#### Scatter plot between Density and Fracture
```{r}
```

#### Smooth line plot using ggplot
```{r}

```

#### Barplot for Crispy
```{r}

```

#### In-built functions for dataframes
```{r}
# Mean oil percentage across all samples

# Mean-centering of OilPercentage

# Sum of the squared deviation from the mean

# Average of the squared deviation from the mean

# Variance of OilPercentage

# Standard deviation of OilPercentage
```


#### Box plot using ggplot with color coding
```{r}

```

















